---
title: "Calculating the Potential for Conflict Index"
output:
  html_document:
    df_print: paged
---

```{r import_raw_data, include=FALSE}
library(here)
library(tidyverse)
library(lubridate)
library(cowplot); theme_set(theme_cowplot())

load(here::here("flow_pref","survey-analysis","flow-pref-analysis-workspace.Rdata"))
```

- The Potential for Conflict Index describes the ammount of agreement or disagreement on flow preferenc among survey respondants.
- Larger PCI scores indicate more disagreement. Smaller PCI scores indicate less disagreement (more agreement)
- For example a PCI score of 1 indicates that respondents are evenly split into two polar-opposit camps (e.g unfavorable v. favorable)
- On the other hand a PCI score of 0 indicates complete consensus among respondents
- This notebook breaks down how PCI is calculated in `flow-pref-analysis.R`

## Step 1: Import tidy survey response data
- Tidy survey response data exported from `tidy-response-data.R` as .Rdata files
- One file contains respondent attribute data
- Another file contains flow preference data
```{r}
load(file = here::here("flow_pref","survey-analysis","respondend-attributes.Rdata"))
load(file = here::here("flow_pref","survey-analysis","flow-pref-data.Rdata"))
```

## Step 2: Calculate average preference and PCI
- Calculate average preference score for each flow bin
- PCI is observed sum of the preference score distance vector, divided by the theoretical maxumum sum of the distance vector
- the `dist` command is used to calculate a distance matrix, quantifying the distance between each pair of preference scores in each flow bin

```{r}
segment_name = "Filter Plant"

# create a new data frame by piping together several commands
results <- flowpref.dat %>%
  
  # Remove NAs and select a specific segment
  filter(is.na(preference.code) == 0 & segment.name == segment_name) %>%
  
  # calculate the average pref score and PCI2 statistic of each flow bin
  group_by(as.numeric(flow)) %>%
  summarize(pref.average = mean(preference.code),   # average preference
            d = sum(abs(dist(preference.code))),    # sum of the score distance vector
            m = sum(abs(dist(rep(c(-3,3),n())))),   # maximum possible sum of distance vector
            pci2 = d/m) %>%                         # PCI2 = d/m
  
  # rename a variable
  rename(flow = `as.numeric(flow)`)

head(results, n = 20)
```

## Step 3: Visualize the results
- Plot avergae flow preference score as a function of flow, marker size is proportional to PCI
- Jitter individual responses as another way of showing respondent agreement/disagreement
```{r}
p1 <- ggplot() +
  geom_point(data = results, aes(x = flow, y = pref.average, size = pci2), color = 'blue') +
  scale_size(range = c(0.1, 5)) +
  geom_jitter(data = filter(flowpref.dat, segment.name == segment_name), aes(x = as.numeric(flow), y = preference.code), size = 1, alpha = 0.3) +
  labs(x = "Flow (cfs)",
       y = "Preference Score") +
  geom_hline(yintercept = 0) +
  ggtitle("Filter Plant flow preference")

print(p1)
```

